---
title: "Matrix from scratch"
author: "Josue Vargas"
date: "2/21/2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Create a vector.
```{r sales_vec}
sales_vec <- c(1000, 1300, 1300, 1100, 1400, 800, 1200, 1500, 1850, 2330, 860, 1300, 1400, 1600, 1970, 570, 380, 450, 465, 580, 155, 190, 210, 250, 300)
```

## Create a matrix from a vector.
```{r vec_to_mtr}
sales_mtr <- matrix(sales_vec, nrow = 5, byrow = TRUE)
```

## Add column and row names to the matrix
```{r add_colnames}
colnames(sales_mtr) <- c(1990, 1991, 1992, 1993, 1994)
```

```{r add_rownames}
rownames(sales_mtr) <- c("Commercial Satellites Delivered",
                         "Satellite Services",
                         "Satellite Ground Equipment",
                         "Commercial Launches",
                         "Remote Sensing Data")
```

Why does the following code produce an error:
`sales_tbl <- rownames_to_column(sales_mtr, var="Industry")`?

The simple answer is `rownames_to_columns()` is a function that only works with data frames. A conceptual answer is
that you are trying to add a `chr` column to a `numerical` matrix. By definition, a matrix can only consist of one data type. In this case: numbers. To use `rownames_to_columns` the matrix must first be converted to a data frame to allow for columns of different data types.

## Convert matrix to a data frame
```{r matrix_to_tibble}
sales_df <- data.frame(sales_mtr, check.names = FALSE)
sales_df <- rownames_to_column(sales_df, var="Industry")
# cannot convert matrix directly to tibble because as_tibble will drop the row names.
sales_tib <- as_tibble(sales_df)
```

## Save data frame to a csv file
```{r saving_data}
write_csv(x = sales_tib, file = "~/Dropbox/programming/R/r_programming/01_data_files/space_industry.csv", col_names = TRUE)
```